---
title: "Untitled"
author: "Troy Liu"
date: "09/16/2025"
output: html_document
---

# Git and GitHub

#### 1) Provide the link to the GitHub repo that you used to practice git from Week 1. It should have:

##### • Your name on the README file.

##### • At least one commit with your name, with a description of what you did in that commit.

-   <https://github.com/zzeng05/Zeng1-Liu2-a1.git>

# Reading Data

#### 2) Read in the .dta version and store in an object called angell_stata.

```{r}
library(haven)
angell_stata <- read_dta("/Users/troyliu/Documents/Education/University of Michigan/Master's Degree/SurvMeth 727/assignment/assignment 1/angell.dta")
angell_stata
```

#### 3) Read in the .txt version and store it in an object called angell_txt.

```{r}
angell_txt <- read.table("/Users/troyliu/Documents/Education/University of Michigan/Master's Degree/SurvMeth 727/assignment/assignment 1/angell.txt")
angell_txt
```

#### 4) What are the differences between angell_stata and angell_txt? Are there differences in the classes of the individual columns?

```{r}
class(angell_stata)
class(angell_txt)
summary(angell_stata)
summary(angell_txt)
lapply(angell_txt, class)
lapply(angell_stata, class)
```

-   At the overall data frame level, angell_txt is simply a base R data frame while angell_stata is stored as a tibble (tidyverse version of the data frame).
-   At variable level, column names are different between angell_stata and angell_txt.
-   At individual column level, There are no differences in the data type of individual columns between them.

#### 5) Make any updates necessary so that angell_txt is the same as angell_stata.

```{r}
# rename angell_txt colnames
colnames(angell_txt) <- colnames(angell_stata)
# re-run to check txt result
summary(angell_txt)
summary(angell_stata)
```

#### 6) Describe the Ethnic Heterogeneity variable. Use descriptive statistics such as mean, median, standard deviation, etc. How does it differ by region?

```{r}
# Descriptive statistics of ethhet
library(tidyverse)
angell_stata %>% 
  summarise(mean_ethhet = mean(ethhet),
            median_ethhet = median(ethhet),
            sd_ethhet = sd(ethhet))

# Descriptive statistics of ethhet by region
angell_stata %>% 
  group_by(region) %>% 
  summarise(mean_ethhet = mean(ethhet),
            median_ethhet = median(ethhet),
            sd_ethhet = sd(ethhet))
```

-   The average non-white and foreign-born white residents among cities is about 31.37%, with a median of 23.7% and a standard deviation of 20.41%.
-   The South region has the most ethnic heterogeneity with the highest mean and spread (through standard deviation); The West has the least ethnic heterogeneity. The Mid-West and East have the moderate level of diversity while the East is slightly higher than the Mid-West.

# Describing Data

#### 7) Install the “MASS” package, load the package. Then, load the Boston dataset.

```{r}
library(MASS)
help(package = "MASS")
data("Boston")
summary(Boston)
head(Boston)
```

#### 8) What is the type of the Boston object?

```{r}
typeof(Boston)
```

-   It's a list.

#### 9) What is the class of the Boston object?

```{r}
class(Boston)
```

-   It's a data frame.

#### 10) How many of the suburbs in the Boston data set bound the Charles river?

```{r}
length(which(Boston$chas == 1))
```

-   There are 35 suburbs bound the Charles river.

#### 11) Do any of the suburbs of Boston appear to have particularly high crime rates? Tax rates? Pupil-teacher ratios? Comment on the range of each variable.

```{r}
# highest values of each variable
BOS <- Boston %>% rownames_to_column("suburb_name")
## crime rates
BOS %>% 
  slice_max(crim)
##tax rates
BOS %>% 
  slice_max(tax)
##pupil-teacher ratios
BOS %>% 
  slice_max(ptratio)

# range of each variable
range(Boston$crim)
range(Boston$tax)
range(Boston$ptratio)
```

-   The #381 has the highest crime rates; #489-493 (5 in total) have the highest tax rates; #355 and #356 has the highest pupil-teacher ratios.
-   The range of crime rates is between .0063 and 88.98; Range of tax rates is between 187 and 711; The range of pupil-teacher ratios is between 12.6 and 22.

#### 12) Describe the distribution of pupil-teacher ratio among the towns in this data set that have a per capita crime rate larger than 1. How does it differ from towns that have a per capita crime rate smaller than 1?

```{r}
# categorizing >1 | <1
h_crim <- subset(Boston, crim > 1)
l_crim <- subset(Boston, crim < 1)
# summary statistics
summary_crim <- list(
  "Crime Rate > 1" = summary(h_crim$ptratio),
  "Crime Rate < 1" = summary(l_crim$ptratio)
)

summary_crim

boxplot(h_crim$ptratio, l_crim$ptratio,
        names = c("Crime Rate > 1", "Crime Rate < 1"),
        main = "Pupil-Teacher Ratio by Crime Rate Group",
        ylab = "Pupil-Teacher Ratio",
        col = c("red", "green"))
```

-   Towns with crime rate greater than 1 tend to have higher pupil-teacher ratios compared to towns with a crime rate less than 1; High-crime towns show both higher mean and median ratios, with more concentrated distribution than low-crime towns.

# Writing Functions

#### 13) Write a function that calculates 95% confidence intervals for a point estimate. The function should be called my_CI. When called with my_CI(2, 0.2), the function should print out “The 95% CI upper bound of point estimate 2 with standard error 0.2 is 2.392. The lower bound is 1.608.”

##### Note: The function should take a point estimate and its standard error as arguments. You may use the formula for 95% CI: point estimate +/- 1.96\*standard error. Hint: Pasting text in R can be done with: paste() and paste0()

```{r}
my_CI <- function(pe, se) {
  
  z <- 1.96  
  up <- pe + z * se
  lo <- pe - z * se
  
  cat(paste0("The 95% CI upper bound of point estimate ", pe, 
      " with standard error ", se, 
      " is ", round(up, 3), 
      ". The lower bound is ", round(lo, 3), ".\n"))
  
}

# testing
my_CI(2, .2)
```

#### 14) Create a new function called my_CI2 that does that same thing as the my_CI function but outputs a vector of length 2 with the lower and upper bound of the confidence interval instead of printing out the text. Use this to find the 95% confidence interval for a point estimate of 0 and standard error 0.4.

```{r}
my_CI2 <- function(pe, se) {
  
  z <- 1.96  
  up <- pe + z * se
  lo <- pe - z * se
  
  return(c(lo, up))
  
}

# testing
my_CI2(0, .4)
```

#### 15) Update the my_CI2 function to take any confidence level instead of only 95%. Call the new function my_CI3. You should add an argument to your function for confidence level. Hint: Use the qnorm function to find the appropriate z-value. For example, for a 95% confidence interval, using qnorm(0.975) gives approximately 1.96.

```{r}
my_CI3 <- function(pe, se, cl = .95) {

  z <- qnorm(1 - (1 - cl)/2)
  up <- pe + z * se
  lo <- pe - z * se
  
  return(c(lo, up))
  
}

# testing
my_CI3(0, .4, cl = .95)
my_CI3(0, .4, cl = .99)
```

#### 16) Without hardcoding any numbers in the code, find a 99% confidence interval for Ethnic Heterogeneity in the Angell dataset. Find the standard error by dividing the standard deviation by the square root of the sample size.

```{r}
# calculate the n and SE for angell data
angell_n <- length(angell_stata$ethhet)
angell_se <- sd(angell_stata$ethhet) / sqrt(angell_n)

# apply the function
my_CI3(mean(angell_stata$ethhet), 
       angell_se,
      cl = 0.99)
```

#### 17) Write a function that you can apply to the Angell dataset to get 95% confidence intervals. The function should take one argument: a vector. Use if-else statements to output NA and avoid error messages if the column in the data frame is not numeric or logical.

```{r}
angell_CI <- function(x) {

  if (!is.numeric(x) && !is.logical(x)) {
    return(NA)
    
  }
  
  else {
    n <- length(x)
    mean_x <- mean(x, na.rm = TRUE)
    se <- sd(x, na.rm = TRUE) / sqrt(n)
    z <- 1.96
    up <- mean_x + z * se
    lo <- mean_x - z * se

    
    return(c(lo = lo, up = up))
    
  }
}

# testing numeric column
angell_CI(angell_stata$ethhet)

# testing non-numeric column
angell_CI(angell_stata$city)
```
